---
title: "create input files for cellphonedb"
author: ""
date: ""
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## Load packages

```{r load-packages, warning = FALSE}
suppressPackageStartupMessages({
  library(scater)
  library(dplyr)
  library(reshape2)
  library(Matrix)
  library(purrr)
  library(scran)
  library(Seurat)
  library(mclust)
  library(tidyverse)
  library(here)
  library(viridis)
  library(RColorBrewer)
  library(pheatmap)
})
```


## set color Vectors
```{r color vectors}

colBRC <- c("#87c5af", "#94033b", "#fc8f3b", "#2580fe", "#514e72")
names(colBRC) <- c("LZFDC", "MRC", "TBRC","DZFDC","PRC2")


colorSLO <- c(viridis(3)[c(1,2)], brewer.pal(n=9, name="YlOrRd")[c(4)])
colorCond <- brewer.pal(n=3, name="Paired")[c(1,2)]
colorSLOCond <- c("#440154FF","#807DBA", "#FEB24C", "#F16913", "#21908CFF")

names(colorSLO) <- c("LN", "PP", "Spleen")
names(colorCond) <- c("naive", "immunized")
names(colorSLOCond) <- c("LN_immunized","LN_naive", "Spleen_naive",
                         "Spleen_immunized", "PP_immunized")


colLYM <- c("#a0b9c4", "#78abb7", "#567a83", "#3c454a", "#0C3C52",
            "#924440", "#eab562",
            "#f1d75f", "#b5a147", "#786c30")


names(colLYM) <- c("naiveBcells", "LZBcells", "DZshmBcells", "proDZ",
                   "plasmaCells", 
                   "Tfh", "DCs",
                   "mdMacrophages", "trMacrophages", "infMacrophages")

colLYM2 <- c("#a0b9c4", "#567a83", "#0C3C52",
            "#924440", "#eab562",
            "#b5a147")


names(colLYM2) <- c("naiveBcells", "GCBcells",
                   "plasmaCells", 
                   "Tfh", "DCs",
                   "Myeloids")

```


## load input data
```{r input data}
basedir <- here()
seuratLY <- readRDS(file = paste0(basedir, 
                              "/data/lympho/allSamples_mergedLab_seurat.rds"))
seuratLY <- subset(seuratLY, clusterLabel %in% names(colLYM))

GCBs <- c("LZBcells", "DZshmBcells", "proDZ")
Myel <- c("mdMacrophages", "trMacrophages", "infMacrophages")
seuratLY$clusterLabel2 <- seuratLY$clusterLabel
seuratLY$clusterLabel2[which(seuratLY$clusterLabel %in% GCBs)] <- "GCBcells"
seuratLY$clusterLabel2[which(seuratLY$clusterLabel %in% Myel)] <- "Myeloids"

seuratBRC <- readRDS(file=paste0(basedir, "/data/folBRC_allSLO_seurat.rds"))

## Each SLO individually
seuratLY$label_plus_SLO <- paste0(seuratLY$clusterLabel2, "_", seuratLY$SLO)
seuratBRC$label_plus_SLO <- paste0(seuratBRC$clusterLabel, "_", seuratBRC$SLO)


## plot input data
Idents(seuratLY) <- seuratLY$clusterLabel2
DimPlot(seuratLY, reduction = "umap", cols = colLYM2)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

Idents(seuratBRC) <- seuratBRC$clusterLabel
DimPlot(seuratBRC, reduction = "umap", cols = colBRC)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

## assign groups
```{r assign groups}

seuratBRC$group <- "BRC"
seuratLY$group <- "LYMPHO"

Idents(seuratLY) <- seuratLY$label_plus_SLO
Idents(seuratBRC) <- seuratBRC$label_plus_SLO

```


## read sel genes
```{r read sel genes}

genes <- data.frame(gene=rownames(seuratBRC)) %>% 
  mutate(geneID=gsub(".*\\.", "", gene))
selGenes <- read_tsv(file=paste0(basedir, 
                                 "/data/GSEA/subsetSpecificFunctGenes", 
                                 "InteractionSignaling.txt")) %>% 
  left_join(., genes, by="geneID")

genesBRC <- selGenes %>% filter(group=="BRC")
genesLym <- selGenes %>% filter(group=="lympho")

```


## plot heatmap sel genes {.tabset}

```{r avg heatmap funct}

avgHeatmap <- function(seurat, selGenes, scaleDim = "row", colVecIdent, 
                       colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  ## format gene names (depends on how gene list is inputed)
  selGenes <- selGenes %>% 
      mutate(geneID = (str_split(gene, '\\.', simplify = T)[,2]))
  #selGenes <- selGenes$labelNam
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>%
    filter(geneID %in% selGenes$geneID)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes$geneID,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale=scaleDim ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc, breaks = seq(from=-2, to=2,by=0.1),
         color = colorRampPalette(c("#111320", "#4d5374", "#899bff"))(40),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}



```



### BRCs
```{r heatmap BRCs, fig.height=9, fig.width=7}

gapC <- seq(from = 3, to = length(levels(seuratBRC)), by = 3)
grpCnt <- genesBRC %>% group_by(functGrp) %>% summarise(cnt=n())
gapR <- data.frame(functGrp=unique(genesBRC$functGrp)) %>% 
  left_join(.,grpCnt, by="functGrp") %>% mutate(cumSum=cumsum(cnt)) 


ordBRC <- c("LZFDC", "DZFDC", "MRC", "TBRC", "PRC2")
ordSLO <- c("LN", "Spleen", "PP")
ordDat <- expand.grid(SLO = ordSLO, BRC = ordBRC) %>% 
  mutate(ordVec=paste0(BRC, "_", SLO))

# pHeat <- lapply(unique(genesBRC$functGrp), function(grp){
#   selGenesFunct <- genesBRC %>% filter(functGrp == grp)
#   pOut <- avgHeatmap(seurat = seuratBRC, selGenes = selGenesFunct,
#                   colVecIdent = colBRC, 
#                   ordVec=ordDat$ordVec,
#                   gapVecC=gapC,cc=F, cr=F,
#                   condCol=T, colVecCond = colorSLO)
# })


pOut <- avgHeatmap(seurat = seuratBRC, selGenes = genesBRC,
                  colVecIdent = colBRC, 
                  ordVec=ordDat$ordVec,
                  gapVecC=gapC,cc=F,
                  gapVecR=gapR$cumSum, cr=F,
                  condCol=T, colVecCond = colorSLO)

Idents(seuratBRC) <- seuratBRC$clusterLabel
pOut <- avgHeatmap(seurat = seuratBRC, selGenes = genesBRC,
                  colVecIdent = colBRC, 
                  ordVec=ordBRC,
                  gapVecC=NULL,cc=F,
                  gapVecR=gapR$cumSum, cr=F,
                  condCol=F)

```


### immune cells
```{r heatmap ImmuneCells, fig.height=8, fig.width=8}

gapC <- seq(from = 3, to = length(levels(seuratLY)), by = 3)
grpCnt <- genesLym %>% group_by(functGrp) %>% summarise(cnt=n())
gapR <- data.frame(functGrp=unique(genesLym$functGrp)) %>% 
  left_join(.,grpCnt, by="functGrp") %>% mutate(cumSum=cumsum(cnt)) 


ordLYM <- c("naiveBcells", "GCBcells", "plasmaCells", 
                   "Tfh", "Myeloids", "DCs")
ordSLO <- c("LN", "Spleen", "PP")
ordDat <- expand.grid(SLO = ordSLO, LYM = ordLYM) %>% 
  mutate(ordVec=paste0(LYM, "_", SLO))

# pHeat <- lapply(unique(genesLym$functGrp), function(grp){
#   selGenesFunct <- genesLym %>% filter(functGrp == grp)
#   pOut <- avgHeatmap(seurat = seuratLY, selGenes = selGenesFunct,
#                   colVecIdent = colLYM2, 
#                   ordVec=ordDat$ordVec,
#                   gapVecC=gapC,cc=F, cr=F,
#                   condCol=T, colVecCond = colorSLO)
# })


pOut <- avgHeatmap(seurat = seuratLY, selGenes = genesLym,
                  colVecIdent = colLYM2, 
                  ordVec=ordDat$ordVec,
                  gapVecC=gapC,cc=F,
                  gapVecR=gapR$cumSum, cr=F,
                  condCol=T, colVecCond = colorSLO)


Idents(seuratLY) <- seuratLY$clusterLabel2
pOut <- avgHeatmap(seurat = seuratLY, selGenes = genesLym,
                  colVecIdent = colLYM2, 
                  ordVec=ordLYM,
                  gapVecC=NULL,cc=F,
                  gapVecR=gapR$cumSum, cr=F,
                  condCol=F)

```

## violin plot sel genes {.tabset}
### BRC
```{r violin sel genes BRC}

seuratBRC$clusterLabel <- factor(seuratBRC$clusterLabel, levels = ordBRC)
pList <- sapply(genesBRC$gene, function(x){
  p <- VlnPlot(object = seuratBRC, features = x,
               group.by = "clusterLabel",
               cols = colBRC, pt.size = 0
               )
  plot(p)
})

```

### Immune cells
```{r violin sel genes ImmCells}

seuratLY$clusterLabel2 <- factor(seuratLY$clusterLabel2, levels = ordLYM)
pList <- sapply(genesLym$gene, function(x){
  p <- VlnPlot(object = seuratLY, features = x,
               group.by = "clusterLabel2",
               cols = colLYM2, pt.size = 0
               )
  plot(p)
})

```


## downstream signaling
```{r}
genes <- data.frame(gene=rownames(seuratBRC)) %>% 
  mutate(geneID=gsub(".*\\.", "", gene))
selGenes <- read_tsv(file=paste0(basedir, 
                            "/data/cellphonedb/downstreamSignalingPW.txt")) %>% 
  left_join(., genes, by="geneID") 

```

### vlnPlots {.tabset}
```{r plot vlnPlots dowstream pw}

allSign <- unique(selGenes$signaling)
template_vp <- c(
    "#### {{sig}}\n",
    "```{r down vp {{sig}}}\n",
    "sign <- as.character('{{sig}}')",
    "signGenes <- selGenes %>% dplyr::filter(signaling == sign)",
    "pList <- sapply(signGenes$gene, function(x){
        p <- VlnPlot(object = seuratBRC, features = x,
               group.by = 'clusterLabel',
               cols = colBRC, pt.size = 0
               )
        plot(p)
        })",
    "```\n",
    "\n"
)
    
plots_vp <- lapply(allSign, 
  function(sig) knitr::knit_expand(text = template_vp)
)

```

`r knitr::knit(text = unlist(plots_vp))`


## session info
```{r session info}
sessionInfo()
date()
```

