---
title: "process cellPhone ouput imm BRC - LYMPHO"
author: ""
date: ""
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## Load packages

```{r load-packages, warning = FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(corrplot)
  library(SummarizedExperiment)
  library(RColorBrewer)
  library(here)
  library(ggsci)
  library(pheatmap)
  library(ggplot2)
  library(reshape2)
  library(viridis)
  library(circlize)
  library(ggpubr)
})
```

## set parameter
```{r read input wt}
basedir <- paste0(here(), "/data/")

colorSLO <- c(viridis(3)[c(1,2)], brewer.pal(n=9, name="YlOrRd")[c(4)])
names(colorSLO) <- c("LN", "PP", "SP")

colBRC <- c("#87c5af", "#94033b", "#fc8f3b", "#2580fe", "#514e72")
names(colBRC) <- c("LZFDC", "MRC", "TBRC","DZFDC","PRC2")

# ## load selected intPairs
# selIntPairs <- read_tsv(paste0(basedir, "/cellphonedb/selIntPairs2.txt"),
#                         col_names = T)

colLab <- c("#a0b9c4", "#567a83", "#0C3C52",
            "#924440", "#eab562",
            "#b5a147")
names(colLab) <- c("naiveBcells", "GCBcells",
                   "plasmaCells", 
                   "Tfh", "DCs",
                   "Myeloids")


ordLymph <- c("LYMPHO_naiveBcells", "LYMPHO_GCBcells", "LYMPHO_plasmaCells",
              "LYMPHO_Tfh", "LYMPHO_Myeloids", "LYMPHO_DCs")
ordLymphShort <- gsub("LYMPHO_", "", ordLymph)

ordBRC <- c("LZFDC", "DZFDC", "MRC", "TBRC", "PRC2")



```


## count significant interactions CellPhoneDB {.tabset}

### format data
```{r format data wt}

sloVec <- c("LN", "PP", "SP")

intList <- lapply(sloVec, function(slo){
  means <- read_tsv(paste0(basedir, "cellphonedb/out", slo, "immGrpNO/means.txt"))
  pVal <- read_tsv(paste0(basedir, "cellphonedb/out", slo,"immGrpNO/pvalues.txt"))

  ### create intPairs
  #### means
  meansCut  <- means %>%
    dplyr::select(-c(id_cp_interaction, partner_a, partner_b, gene_a, gene_b,
              receptor_a, receptor_b, annotation_strategy, secreted, is_integrin,
              interacting_pair))
  rownames(meansCut) <- means$interacting_pair

  #### pVal
  pValCut  <- pVal %>%
    dplyr::select(-c(id_cp_interaction, partner_a, partner_b, gene_a, gene_b,
              receptor_a, receptor_b, annotation_strategy, secreted, is_integrin,
              interacting_pair))
  rownames(pValCut) <- pVal$interacting_pair
  
  #### cellA/cellB
  CellNames <- data.frame(CellsInt = as.character(colnames(meansCut))) %>%
    mutate(cellA=gsub("\\|.*","" ,CellsInt)) %>%
    mutate(cellB=gsub("^.*\\|","" ,CellsInt)) %>% 
    mutate(groupA=gsub("_.*","" ,CellsInt)) %>% 
    mutate(groupB=gsub("_.*","",gsub("^.*\\|","" ,CellsInt))) 


  ### summarize mean
  meansCut_long <- meansCut %>% rownames_to_column(var="intPair") %>%
    melt(., id.vars=c("intPair"), measure.vars=colnames(meansCut),
         variable.name="CellsInt", value.name="mean") %>% 
    left_join(., CellNames, by="CellsInt") %>% 
    filter(!groupA == groupB)

  ### summarize pVal
  pValCut_long <- pValCut %>% rownames_to_column(var="intPair") %>%
    melt(., id.vars=c("intPair"), measure.vars=colnames(meansCut),
         variable.name="CellsInt", value.name="pVal") %>% 
    left_join(., CellNames, by="CellsInt") %>% 
    filter(!groupA == groupB)

  allSum <- meansCut_long %>% mutate(pVal=pValCut_long$pVal) %>%
    mutate(SLO = slo) %>% 
    mutate(sign=ifelse(pVal<0.05, "p < 0.05", "p > 0.05"))

})
names(intList) <- sloVec


```


### all sign int {.tabset}
#### heatmap
```{r all sig int heatmap, fig.height=6, fig.width=6}

## filter for sign int
signInt <- lapply(intList, function(all){
  allSig <- all %>% filter(pVal<0.05) 
})
names(signInt) <- sloVec

signIntDat <- data.frame(do.call("rbind", signInt))

## reorder int to always BRC to LYMPHO and remove selPop 
reordInt <- signIntDat %>% filter(groupA=="LYMPHO") %>% 
  mutate(intA=gsub("_.*", "", intPair)) %>% 
  mutate(intB=gsub(".*_", "", intPair))
reordIntFin <- data.frame(intPair=paste0(reordInt$intB, "_", reordInt$intA),
                          CellsInt=paste0(reordInt$cellB, "|", reordInt$cellA),
                          mean=reordInt$mean,
                          cellA=reordInt$cellB,
                          cellB=reordInt$cellA,
                          groupA=reordInt$groupB,
                          groupB=reordInt$groupA,
                          pVal=reordInt$pVal,
                          SLO=reordInt$SLO,
                          sign=reordInt$sign)

signIntDatOrd <- rbind(signIntDat[which(signIntDat$groupA=="BRC"),],
                       reordIntFin) 

### -------------------- all sign INT -------------------------------------- ###
## cnt sign int for each SLO for each intPair + cnts rel to lymph population
lymphCntTot <- signIntDatOrd %>% group_by(cellB, SLO) %>% 
  summarize(cntTotL=n())

cntPairSLO <- signIntDatOrd %>% group_by(cellA, cellB, SLO) %>% 
  summarize(cntPair=n()) %>%
  left_join(., lymphCntTot, by = c("cellB", "SLO")) %>% 
  mutate(relCnt=cntPair/cntTotL)

## heatmap with rel cnts per SLO
cntPairSLOform <- cntPairSLO %>%
  mutate(BRC=paste0(SLO, "_", gsub(".*_", "", cellA))) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, cntPair) %>% 
  spread(., BRC, cntPair)
cntPairSLOmat <- cntPairSLOform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairSLOmat) <- cntPairSLOform$LYMPHO

annotation_col <-data.frame(clu=gsub("(^.*?_)","", colnames(cntPairSLOmat))) %>%
    dplyr::mutate(SLO=gsub("(_.*$)","",colnames(cntPairSLOmat))) 
rownames(annotation_col) <- colnames(cntPairSLOmat)  
ann_colors <- list(clu = colBRC, 
                   SLO=colorSLO)
gapVecC <- c(5,10)

#### scale across all SLO
pheatmap(cntPairSLOmat, scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = gapVecC)

#### each SLO indvidually
lapply(sloVec, function(slo){
  cntPairSLOmatSub <- cntPairSLOmat[,grepl(paste0(slo, "_*"),
                                           colnames(cntPairSLOmat))]

  pheatmap(cntPairSLOmatSub, scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)
  out <- NULL
})


```

#### circle plots for each SLO
```{r circle plot all sign for each SLO}

### relative cnts
adjListConsInt <- cntPairSLO  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(BRCsub, IMMsub, SLO, relCnt) %>% 
  arrange(IMMsub) %>% 
  arrange(BRCsub)

lapply(sloVec, function(slo){
  adjSub <- adjListConsInt %>% filter(SLO==slo) %>% 
    dplyr::select(-SLO)
  
  chordDiagram(adjSub, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
  circos.clear()
  
})


### absolute cnts
adjListConsInt <- cntPairSLO  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(BRCsub, IMMsub, SLO, cntPair) %>% 
  arrange(IMMsub) %>% 
  arrange(BRCsub)

lapply(sloVec, function(slo){
  adjSub <- adjListConsInt %>% filter(SLO==slo) %>% 
    dplyr::select(-SLO)
  
  chordDiagram(adjSub, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
  circos.clear()
  
})


```



### conserved sign int {.tabset}
#### heatmap cons all
```{r cons sign int all}

### ------------------ conserved sign INT ---------------------------------- ###

## interactions conserved across all 3 SLOs
signIntCons <- signIntDatOrd %>% 
  dplyr::select(intPair, cellA, cellB, groupA, groupB) %>% 
  group_by(intPair, cellA, cellB, groupA, groupB) %>%
  summarise(cnt=n()) %>% 
  filter(cnt >2)

## write file
signIntConsOut <- signIntCons %>% ungroup(.) %>% 
  dplyr::select(intPair, cellA, cellB)
write.table(signIntConsOut, file=paste0(basedir,
                                        "cellphonedb/signIntImmGrp", 
                                        "_conservedSLO.txt"),
            quote = F, col.names = T, row.names = F, sep = "\t")

## cnt conserved sign interactions for each intPair + rel cnt
lymphCntTot <- signIntCons %>% group_by(cellB) %>% 
  summarize(cntTotL=n())
cntPairCons <- signIntCons %>% group_by(cellA, cellB) %>% 
  summarize(cntPair=n()) %>% left_join(., lymphCntTot, by = "cellB") %>% 
  mutate(relCnt=cntPair/cntTotL)
meanBRC <- cntPairCons %>% group_by(cellA) %>% 
  summarise(meanRel=mean(relCnt), sdRel=sd(relCnt), relLyTot = sum(relCnt))
cntPairCons <- cntPairCons %>% left_join(., meanBRC, by= "cellA") %>% 
  mutate(zScore=(relCnt-meanRel)/sdRel) %>% 
  mutate(relRelCnt=relCnt/relLyTot)
  
### heatmap with cons int
cntPairConsform <- cntPairCons %>%
  mutate(BRC=gsub(".*_", "", cellA)) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, cntPair) %>% 
  spread(., BRC, cntPair)
cntPairConsmat <- cntPairConsform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairConsmat) <- cntPairConsform$LYMPHO

annotation_col <-data.frame(lymph=rownames(cntPairConsmat)) 
rownames(annotation_col) <- rownames(cntPairConsmat)  
ann_colors <- list(lymph=colLab)

ordLymph2 <- gsub(".*_", "", ordLymph)
cntPairConsmat <- cntPairConsmat[,ordBRC]
cntPairConsmat <- cntPairConsmat[ordLymph2,]
pheatmap(t(cntPairConsmat), scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)

## per BRC
annotation_col <-data.frame(clu=colnames(cntPairConsmat)) 
rownames(annotation_col) <- colnames(cntPairConsmat)  
ann_colors <- list(clu = colBRC)
pheatmap(cntPairConsmat, scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)

### heatmap with cons int rel cnts..
cntPairConsform <- cntPairCons %>%
  mutate(BRC=gsub(".*_", "", cellA)) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, relCnt) %>% 
  spread(., BRC, relCnt)
cntPairConsmat <- cntPairConsform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairConsmat) <- cntPairConsform$LYMPHO

annotation_col <-data.frame(lymph=rownames(cntPairConsmat)) 
rownames(annotation_col) <- rownames(cntPairConsmat)  
ann_colors <- list(lymph=colLab)

cntPairConsmat <- cntPairConsmat[,ordBRC]
cntPairConsmat <- cntPairConsmat[ordLymph2,]

pheatmap(t(cntPairConsmat), scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#7583bf", "#9cafff"))(20),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)


### barplot
cntPairConsBar <- cntPairCons %>%
  mutate(ImmuneCell=gsub("LYMPHO_", "", cellB)) %>% 
  mutate(BRC=gsub("BRC_", "", cellA)) 
  
p <- ggbarplot(cntPairConsBar, x = "BRC", y = "relRelCnt",
               color = "ImmuneCell", fill= "ImmuneCell",
               palette = colLab)
p

### Dotplot
p <- ggplot(cntPairConsBar, aes(x=BRC,y=ImmuneCell)) +
  geom_point(aes(size=relRelCnt,color=ImmuneCell)) +
  scale_color_manual(values=colLab) +
  scale_radius(range = c(0, 8)) +
  scale_y_discrete(limits=ordLymphShort) +
  scale_x_discrete(limits=ordBRC) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text=element_text(size=10, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title=element_blank(),
        panel.border = element_rect(size = 0.7, linetype = "solid",
                                    colour = "black"))
p

```


#### heatmap all int independent of conservation
```{r sign int all no cons dotplot, fig.height=3, fig.width=4.5}

### ------------------ conserved sign INT ---------------------------------- ###

## all interactions independent of conservation
signIntCons <- signIntDatOrd %>% 
  dplyr::select(intPair, cellA, cellB, groupA, groupB) %>% 
  group_by(intPair, cellA, cellB, groupA, groupB) %>%
  summarise(cnt=n()) %>% 
  filter(cnt >0)


## cnt conserved sign interactions for each intPair + rel cnt
lymphCntTot <- signIntCons %>% group_by(cellB) %>% 
  summarize(cntTotL=n())
cntPairCons <- signIntCons %>% group_by(cellA, cellB) %>% 
  summarize(cntPair=n()) %>% left_join(., lymphCntTot, by = "cellB") %>% 
  mutate(relCnt=cntPair/cntTotL)
meanBRC <- cntPairCons %>% group_by(cellA) %>% 
  summarise(meanRel=mean(relCnt), sdRel=sd(relCnt), relLyTot = sum(relCnt))
cntPairCons <- cntPairCons %>% left_join(., meanBRC, by= "cellA") %>% 
  mutate(zScore=(relCnt-meanRel)/sdRel) %>% 
  mutate(relRelCnt=relCnt/relLyTot)
  
### heatmap with cons int
cntPairConsform <- cntPairCons %>%
  mutate(BRC=gsub(".*_", "", cellA)) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, cntPair) %>% 
  spread(., BRC, cntPair)
cntPairConsmat <- cntPairConsform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairConsmat) <- cntPairConsform$LYMPHO

annotation_col <-data.frame(lymph=rownames(cntPairConsmat)) 
rownames(annotation_col) <- rownames(cntPairConsmat)  
ann_colors <- list(lymph=colLab)

ordLymph2 <- gsub(".*_", "", ordLymph)
cntPairConsmat <- cntPairConsmat[,ordBRC]
cntPairConsmat <- cntPairConsmat[ordLymph2,]
pheatmap(t(cntPairConsmat), scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)

## per BRC
annotation_col <-data.frame(clu=colnames(cntPairConsmat)) 
rownames(annotation_col) <- colnames(cntPairConsmat)  
ann_colors <- list(clu = colBRC)
pheatmap(cntPairConsmat, scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)

### heatmap with cons int rel cnts..
cntPairConsform <- cntPairCons %>%
  mutate(BRC=gsub(".*_", "", cellA)) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, relCnt) %>% 
  spread(., BRC, relCnt)
cntPairConsmat <- cntPairConsform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairConsmat) <- cntPairConsform$LYMPHO

annotation_col <-data.frame(lymph=rownames(cntPairConsmat)) 
rownames(annotation_col) <- rownames(cntPairConsmat)  
ann_colors <- list(lymph=colLab)

cntPairConsmat <- cntPairConsmat[,ordBRC]
cntPairConsmat <- cntPairConsmat[ordLymph2,]

pheatmap(t(cntPairConsmat), scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#7583bf", "#9cafff"))(20),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)


### barplot
cntPairConsBar <- cntPairCons %>%
  mutate(ImmuneCell=gsub("LYMPHO_", "", cellB)) %>% 
  mutate(BRC=gsub("BRC_", "", cellA)) 
  
p <- ggbarplot(cntPairConsBar, x = "BRC", y = "relRelCnt",
               color = "ImmuneCell", fill= "ImmuneCell",
               palette = colLab)
p

### Dotplot
p <- ggplot(cntPairConsBar, aes(x=BRC,y=ImmuneCell)) +
  geom_point(aes(size=relRelCnt,color=ImmuneCell)) +
  scale_color_manual(values=colLab) +
  scale_radius(range = c(1, 7)) +
  scale_y_discrete(limits=ordLymphShort) +
  scale_x_discrete(limits=ordBRC) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text=element_text(size=10, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title=element_blank(),
        panel.border = element_rect(size = 0.7, linetype = "solid",
                                    colour = "black"))
p

```



#### heatmap cons min 2
```{r cons sign int min 2, fig.height=4, fig.width=6}

### ------------------ conserved sign INT ---------------------------------- ###

## interactions conserved in min 2 SLOs
signIntCons <- signIntDatOrd %>% 
  dplyr::select(intPair, cellA, cellB, groupA, groupB) %>% 
  group_by(intPair, cellA, cellB, groupA, groupB) %>%
  summarise(cnt=n()) %>% 
  filter(cnt >1)


## cnt conserved sign interactions for each intPair + rel cnt
lymphCntTot <- signIntCons %>% group_by(cellB) %>% 
  summarize(cntTotL=n())
cntPairCons <- signIntCons %>% group_by(cellA, cellB) %>% 
  summarize(cntPair=n()) %>% left_join(., lymphCntTot, by = "cellB") %>% 
  mutate(relCnt=cntPair/cntTotL)
meanBRC <- cntPairCons %>% group_by(cellA) %>% 
  summarise(meanRel=mean(relCnt), sdRel=sd(relCnt), relLyTot = sum(relCnt))
cntPairCons <- cntPairCons %>% left_join(., meanBRC, by= "cellA") %>% 
  mutate(zScore=(relCnt-meanRel)/sdRel) %>% 
  mutate(relRelCnt=relCnt/relLyTot)
  
### heatmap with cons int
cntPairConsform <- cntPairCons %>%
  mutate(BRC=gsub(".*_", "", cellA)) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, cntPair) %>% 
  spread(., BRC, cntPair)
cntPairConsmat <- cntPairConsform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairConsmat) <- cntPairConsform$LYMPHO

annotation_col <-data.frame(lymph=rownames(cntPairConsmat)) 
rownames(annotation_col) <- rownames(cntPairConsmat)  
ann_colors <- list(lymph=colLab)

ordLymph2 <- gsub(".*_", "", ordLymph)
cntPairConsmat <- cntPairConsmat[,ordBRC]
cntPairConsmat <- cntPairConsmat[ordLymph2,]
pheatmap(t(cntPairConsmat), scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)

## per BRC
annotation_col <-data.frame(clu=colnames(cntPairConsmat)) 
rownames(annotation_col) <- colnames(cntPairConsmat)  
ann_colors <- list(clu = colBRC)
pheatmap(cntPairConsmat, scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#4e5780", "#9cafff"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)

### heatmap with cons int rel cnts..
cntPairConsform <- cntPairCons %>%
  mutate(BRC=gsub(".*_", "", cellA)) %>% 
  mutate(LYMPHO=gsub(".*_", "", cellB)) %>% ungroup(.) %>% 
  dplyr::select(BRC, LYMPHO, relCnt) %>% 
  spread(., BRC, relCnt)
cntPairConsmat <- cntPairConsform %>% dplyr::select(-LYMPHO) %>% 
  as.matrix(.)
rownames(cntPairConsmat) <- cntPairConsform$LYMPHO

annotation_col <-data.frame(lymph=rownames(cntPairConsmat)) 
rownames(annotation_col) <- rownames(cntPairConsmat)  
ann_colors <- list(lymph=colLab)

cntPairConsmat <- cntPairConsmat[,ordBRC]
cntPairConsmat <- cntPairConsmat[ordLymph2,]

pheatmap(t(cntPairConsmat), scale="row" ,treeheight_row = 0, cluster_rows = F, 
         cluster_cols = F,
         color = colorRampPalette(c("#11161d", "#7583bf", "#9cafff"))(20),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = NULL, gaps_col = NULL)


### barplot
cntPairConsBar <- cntPairCons %>%
  mutate(ImmuneCell=gsub("LYMPHO_", "", cellB)) %>% 
  mutate(BRC=gsub("BRC_", "", cellA)) 
  
p <- ggbarplot(cntPairConsBar, x = "BRC", y = "relRelCnt",
               color = "ImmuneCell", fill= "ImmuneCell",
               palette = colLab)
p

### Dotplot
p <- ggplot(cntPairConsBar, aes(x=BRC,y=ImmuneCell)) +
  geom_point(aes(size=relRelCnt,color=ImmuneCell)) +
  scale_color_manual(values=colLab) +
  scale_radius(range = c(0, 8)) +
  scale_y_discrete(limits=ordLymphShort) +
  scale_x_discrete(limits=ordBRC) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text=element_text(size=10, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title=element_blank(),
        panel.border = element_rect(size = 0.7, linetype = "solid",
                                    colour = "black"))
p

```



#### circle plots for each subset
```{r circle pplot for each subset}

### relative cnts
adjListConsInt <- cntPairCons  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(IMMsub, BRCsub, relCnt, relRelCnt) %>% 
  arrange(BRCsub) %>% 
  arrange(IMMsub)

lapply(ordBRC, function(BRC){
  adjSub <- adjListConsInt %>% filter(BRCsub==BRC)
  
  chordDiagram(adjSub, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
  circos.clear()
  
})


### absolute cnts
adjListConsInt <- cntPairCons  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(IMMsub, BRCsub, cntPair) %>% 
  arrange(BRCsub) %>% 
  arrange(IMMsub)
  

lapply(ordBRC, function(BRC){
  adjSub <- adjListConsInt %>% filter(BRCsub==BRC)
  
  chordDiagram(adjSub, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
  circos.clear()
  
})


```


#### circle plots across all subsets
```{r circle plot across all subsets}

### relative cnts
adjListConsInt <- cntPairCons  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(BRCsub, IMMsub, relRelCnt, relCnt) %>% 
  arrange(IMMsub) %>% 
  arrange(BRCsub)
  
chordDiagram(adjListConsInt, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
circos.clear()



### absolute cnts
adjListConsInt <- cntPairCons  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(BRCsub, IMMsub, cntPair) %>% 
  arrange(IMMsub) %>% 
  arrange(BRCsub)
  

chordDiagram(adjListConsInt, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
circos.clear()


```


### fraction of all sign int that are conserved {.tabset}
```{r cons fract of all sign}

signIntSum <- signIntDatOrd %>% dplyr::select(intPair, cellA, cellB) %>% 
  distinct() %>%
  left_join(., signIntCons, by=c("intPair", "cellA", "cellB")) %>% 
  mutate(cons=ifelse(is.na(cnt), "noCons", "Cons"))
consIntSum <- signIntSum %>% group_by(cellA, cellB, cons) %>% 
  summarize(cntPair=n()) %>% ungroup() %>% 
  spread(., cons, cntPair) %>% 
  mutate(sumPair=noCons + Cons) %>% 
  mutate(fracCons=Cons/sumPair)

adjListConsInt <- consIntSum  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(IMMsub, BRCsub, fracCons) %>% 
  arrange(BRCsub) %>% 
  arrange(IMMsub)
  
## for each subset
lapply(ordBRC, function(BRC){
  adjSub <- adjListConsInt %>% filter(BRCsub==BRC)
  
  chordDiagram(adjSub, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
  circos.clear()
  
})

## across all
adjListConsInt <- consIntSum  %>% ungroup() %>% 
  mutate(BRCsub=factor(gsub("BRC_", "", cellA), levels = ordBRC)) %>% 
  mutate(IMMsub=factor(gsub("LYMPHO_", "", cellB), levels = ordLymphShort)) %>% 
  dplyr::select(BRCsub, IMMsub, fracCons) %>% 
  arrange(IMMsub) %>% 
  arrange(BRCsub)
  
chordDiagram(adjListConsInt, big.gap = 15, grid.col = c(colBRC, colLab),
               transparency = 0.1, annotationTrack = c("name", "grid"))
circos.clear()


```


## vis conserved int {.tabset}
### dotplot with cons subset specific int
```{r dotplot cons subSpec int, fig.height=9, fig.width=6}

## treat MRC and DZFDC as same for sel
signIntDatOrd_mod <- signIntDatOrd %>% 
  mutate(cellA=ifelse(cellA=="BRC_MRC", "BRC_DZFDC", cellA)) %>% 
  mutate(CellsInt=paste0(cellA, "|", cellB))


## subset specific interactions that conserved across SLOs
ConsIntSubSpec <- signIntDatOrd_mod %>% group_by(intPair, cellB, SLO) %>% 
  summarise(cntIntP=n()) %>% filter(cntIntP == 1) %>% 
  mutate(sel=paste0(intPair, "_", cellB))
ConsIntSubSpecFil <- signIntCons %>%
  mutate(sel=paste0(intPair, "_", cellB)) %>% 
  mutate(sel2=paste0(intPair, "_", cellA)) %>% 
  filter(sel %in% ConsIntSubSpec$sel) 

## filter for sign int
selIntDat <- data.frame(do.call("rbind", intList))

## reorder int to always BRC to LYMPHO and remove selPop 
reordInt <- selIntDat %>% filter(groupA=="LYMPHO") %>% 
  mutate(intA=gsub("_.*", "", intPair)) %>% 
  mutate(intB=gsub(".*_", "", intPair))
reordIntFin <- data.frame(intPair=paste0(reordInt$intB, "_", reordInt$intA),
                          CellsInt=paste0(reordInt$cellB, "|", reordInt$cellA),
                          mean=reordInt$mean,
                          cellA=reordInt$cellB,
                          cellB=reordInt$cellA,
                          groupA=reordInt$groupB,
                          groupB=reordInt$groupA,
                          pVal=reordInt$pVal,
                          SLO=reordInt$SLO,
                          sign=reordInt$sign)

selIntDatOrd <- rbind(selIntDat[which(selIntDat$groupA=="BRC"),],
                       reordIntFin) 

selIntDatOrd_mean <- selIntDatOrd %>%
  filter(intPair %in% ConsIntSubSpecFil$intPair) %>% 
  group_by(intPair, CellsInt, cellA, cellB) %>% 
  summarise(meanSLO=mean(mean), maxPval=max(pVal)) %>% 
  mutate(logPval=-log(maxPval)) %>% 
  mutate(sel2=paste0(intPair, "_", cellA)) %>% 
  filter(sel2 %in% ConsIntSubSpecFil$sel2) %>% 
  mutate(col=factor(gsub(".*_", "", cellA), levels = ordBRC)) %>% 
  arrange(., col)

selIntDatOrd_mean$logPval[which(selIntDatOrd_mean$logPval > 4)] <- 4

                             
p <- ggplot(selIntDatOrd_mean, aes(x=cellB,y=intPair)) +
  geom_point(aes(size=logPval,color=col)) +
  scale_color_manual(values=colBRC) +
  scale_radius(range = c(0, 4)) +
  scale_y_discrete(limits=unique(selIntDatOrd_mean$intPair)) +
  scale_x_discrete(limits=ordLymph) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text=element_text(size=10, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title=element_blank(),
        panel.border = element_rect(size = 0.7, linetype = "solid",
                                    colour = "black"))
p


```


### dotplot with subset specific int conserved in min 2 SLO

```{r dotplot min 2 SLO subSpec int, fig.height=12, fig.width=6}

# ## interactions conserved across all 3 SLOs
# ## for MRC + DZFDC less stringend as we've fewer cells in PP (cut >=2)
# signIntCons <- signIntDatOrd %>% 
#   dplyr::select(intPair, cellA, cellB, groupA, groupB) %>% 
#   group_by(intPair, cellA, cellB, groupA, groupB) %>%
#   summarise(cnt=n()) %>% 
#   filter(ifelse(cellA %in% c("BRC_DZFDC", "BRC_MRC"), cnt>1, cnt == 3))


## interactions conserved across at least 2 SLOs
signIntCons <- signIntDatOrd %>% 
  dplyr::select(intPair, cellA, cellB, groupA, groupB) %>% 
  group_by(intPair, cellA, cellB, groupA, groupB) %>%
  summarise(cnt=n()) %>% 
  filter(cnt > 1)

signIntConsSel <- signIntCons %>% ungroup() %>% 
  group_by(intPair, cellB) %>%
  summarise(cnt=n()) %>% 
  filter(cnt == 1)

signIntConsSpec <- signIntCons %>% ungroup() %>% 
  left_join(., signIntConsSel, by=c("intPair", "cellB")) %>% 
  filter(!is.na(cnt.y)) %>% 
  mutate(sel2=paste0(intPair, "_", cellA))


selIntDatOrd_mean <- selIntDatOrd %>%
  filter(intPair %in% signIntConsSpec$intPair) %>% 
  group_by(intPair, CellsInt, cellA, cellB) %>% 
  summarise(meanSLO=mean(mean), meanPval=mean(pVal)) %>% 
  mutate(logPval=-log(meanPval)) %>% 
  mutate(sel2=paste0(intPair, "_", cellA)) %>% 
  filter(sel2 %in% signIntConsSpec$sel2) %>% 
  mutate(col=factor(gsub(".*_", "", cellA), levels = ordBRC)) %>% 
  arrange(., col)

selIntDatOrd_mean$logPval[which(selIntDatOrd_mean$logPval > 4)] <- 4

## remove dupl intPairs
dupPairSum <- selIntDatOrd_mean %>% group_by(intPair, cellA) %>% 
  summarise(cnt=n())
dupPair <- dupPairSum$intPair[duplicated(dupPairSum$intPair)]
selIntDatOrd_mean <-  selIntDatOrd_mean %>% filter(!intPair %in% dupPair)
                             
p <- ggplot(selIntDatOrd_mean, aes(x=cellB,y=intPair)) +
  geom_point(aes(size=logPval,color=col)) +
  scale_color_manual(values=colBRC) +
  scale_radius(range = c(0, 4)) +
  scale_y_discrete(limits=unique(selIntDatOrd_mean$intPair)) +
  scale_x_discrete(limits=ordLymph) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text=element_text(size=10, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title=element_blank(),
        panel.border = element_rect(size = 0.7, linetype = "solid",
                                    colour = "black"))
p


```


## session info
```{r session info}
sessionInfo()
date()
```


