---
title: "create input files for cellphonedb"
author: ""
date: ""
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

# Load packages

```{r load-packages, warning = FALSE}
suppressPackageStartupMessages({
  library(scater)
  library(ensembldb)
  library(dplyr)
  library(reshape2)
  library(Matrix)
  library(org.Mm.eg.db)
  library(purrr)
  library(scran)
  library(Seurat)
  library(mclust)
  library(tidyverse)
  library(biomaRt)
  library(here)
})
```


## load input data
```{r input data}
basedir <- here()
seuratLY <- readRDS(file = paste0(basedir, 
                              "/data/lympho/allSamples_mergedLab_seurat.rds"))
seuratBRC <- readRDS(file=paste0(basedir, "/data/folBRC_allSLO_seurat.rds"))

## Each SLO individually
selLab <- c("naiveBcells", "LZBcells", "DZshmBcells", "proDZ",
                   "plasmaCells", 
                   "Tfh", "DCs",
                   "mdMacrophages", "trMacrophages", "infMacrophages")
seuratLY <- subset(seuratLY, clusterLabel %in% selLab)
GCBs <- c("LZBcells", "DZshmBcells", "proDZ")
Myel <- c("mdMacrophages", "trMacrophages", "infMacrophages")
seuratLY$clusterLabel2 <- seuratLY$clusterLabel
seuratLY$clusterLabel2[which(seuratLY$clusterLabel %in% GCBs)] <- "GCBcells"
seuratLY$clusterLabel2[which(seuratLY$clusterLabel %in% Myel)] <- "Myeloids"

seuratLY$label_plus_SLO <- paste0(seuratLY$clusterLabel2, "_", seuratLY$SLO)
seuratLY$clusterLabel <- seuratLY$clusterLabel2

### only imm BRCs
seuratBRC <- subset(seuratBRC, cond == "immunized")

## plot input data
Idents(seuratLY) <- seuratLY$clusterLabel2
DimPlot(seuratLY, reduction = "umap")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

Idents(seuratBRC) <- seuratBRC$clusterLabel
DimPlot(seuratBRC, reduction = "umap")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

# filter cells and assign cluster
```{r filter and assign cluster to cells}

seuratBRC$group <- "BRC"
seuratLY$group <- "LYMPHO"

## only cluster XYZ
## seuratLY <- subset(seuratLY, idents="XYZ")

## subsample seuratLY
Idents(seuratLY) <- seuratLY$label_plus_SLO
#seuratLY <- subset(seuratLY, downsample = 500) ## uncomment to subsample in R script

## subsample seuratBRC
Idents(seuratBRC) <- seuratBRC$SLO_plus_clusterlabel
#seuratBRC <- subset(seuratBRC, downsample = 500) ## uncomment to subsample in R script


Idents(seuratLY) <- seuratLY$clusterLabel
Idents(seuratBRC) <- seuratBRC$clusterLabel
  
## merge
### fix assay and gene names
sceBRC <- as.SingleCellExperiment(seuratBRC)
sceLY <- as.SingleCellExperiment(seuratLY)

genesBRC <- data.frame(geneB=rownames(sceBRC)) %>% 
  mutate(ensID=gsub("\\..*","",geneB))
genesLY <- data.frame(geneL=rownames(sceLY)) %>% 
  mutate(ensID=gsub("\\..*","",geneL))
genesBoth <- full_join(genesBRC, genesLY, by="ensID") %>% 
  mutate(gene=ifelse(is.na(geneL), geneB, geneL)) %>% 
  dplyr::select(ensID, gene)
genesBRC <- genesBRC %>% left_join(., genesBoth, by="ensID")
genesLY <- genesLY %>% left_join(., genesBoth, by="ensID")

rownames(sceBRC) <- genesBRC$gene
rownames(sceLY) <- genesLY$gene

seuratBRC <- as.Seurat(sceBRC)
seuratLY <- as.Seurat(sceLY)

seuratBRC@assays$RNA <- seuratBRC@assays$originalexp
seuratLY@assays$originalexp <- seuratLY@assays$RNA

seuratMerge <- merge(seuratBRC, seuratLY)
seuratMerge <- ScaleData(seuratMerge, assay = "RNA")
seuratMerge <- ScaleData(seuratMerge, assay = "originalexp")

remove(seuratBRC)
remove(seuratLY)

sceMerge <- as.SingleCellExperiment(seuratMerge)
(table(sceMerge@colData$group))

## assign celltype
colData(sceMerge)$clusterFinal <- paste0(colData(sceMerge)$group, "_", 
                                         colData(sceMerge)$clusterLabel)
(table(sceMerge@colData$clusterFinal))

```


# create count matrix for cellphonedb
```{r create cnt mat}

## set human ensIDs as rownames
ensIDs <- gsub("\\..*","",rownames(sceMerge))

## map to human orthologs
mart2 <- useMart("ensembl", dataset="hsapiens_gene_ensembl")
mart1 = useMart("ensembl", dataset="mmusculus_gene_ensembl") 

# human / mouse
geneIDS_hum <- getLDS(attributes=c("ensembl_gene_id"),
           filters="ensembl_gene_id", values=ensIDs, mart=mart1,
           attributesL=c("ensembl_gene_id"), martL=mart2)

geneIDsMatch <- as.data.frame(ensIDs) %>% 
  left_join(., geneIDS_hum, by=c("ensIDs"="Gene.stable.ID")) 
geneIDsMatch$Gene.stable.ID.1[
  which(is.na(geneIDsMatch$Gene.stable.ID.1))] <- "unknown"
geneIDsMatch <- geneIDsMatch[!duplicated(geneIDsMatch$ensIDs),]
rowData(sceMerge)$humGene <- geneIDsMatch$Gene.stable.ID.1
dim(sceMerge)
sceMerge <- sceMerge[which(rowData(sceMerge)$humGene != "unknown"),] 
dim(sceMerge)

## filter for SLO and write cnt mat plus metadat
lapply(unique(sceMerge$SLO), function(slo){
  sceSLO <- sceMerge[,which(sceMerge@colData$SLO == slo)]
  cellPhoneSLO <- cbind(rowData(sceSLO)$humGene, as.matrix(logcounts(sceSLO)))
  colnames(cellPhoneSLO)[1] <- "gene"
  write.table(cellPhoneSLO, 
              file = paste0(basedir,
                            "/data/cellphonedb/folBRCimm_LYMGrpNOsub_",slo,
                            "_countMatrix.txt"),
              row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")

  metaDatSLO <- as.data.frame(cbind(colnames(sceSLO),
                                    colData(sceSLO)$clusterFinal))
  colnames(metaDatSLO) <- c("Cell", "cell_type")
  write.table(metaDatSLO, 
              file = paste0(basedir,
                            "/data/cellphonedb/folBRCimm_LYMGrpNOsub_",slo,
                            "_metaData.txt"),
              row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")


})



```


## session info
```{r session info}
sessionInfo()
date()
```



cd /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb
eval "$(pyenv init -)"
eval "$(pyenv virtualenv-init -)"

## run cellphonedb
### build database
cellphonedb database generate --user-interactions custom_interaction_file.csv

##run statistical analysis
## subsampling in R script.. 
cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpNOsub_LN_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpNOsub_LN_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outLNimmGrpNO/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpNOsub_PP_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpNOsub_PP_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outPPimmGrpNO/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpNOsub_Spleen_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpNOsub_Spleen_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outSPimmGrpNO/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100


Count network interactions:
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpsub_LN_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outLNimmGrp/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outLNimmGrp
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpsub_PP_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outPPimmGrp/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outPPimmGrp
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRCimm_LYMGrpsub_Spleen_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outSPimmGrp/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outSPimmGrp

We use CellPhoneDB as tool to predict cellular communication between SILT FRCs and ILC populations in our dataset. In brief, pairs of known interacting ligands and receptors are derived from public databases and cell types are analysed for enriched receptor–ligand interactions based on expression of a receptor by one cell type and a ligand by another cell type. 

deactivate  
