---
title: "create LN reference for spatial decomposition"
author: ""
date: ""
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

# Load packages

```{r load-packages, warning = FALSE}
suppressPackageStartupMessages({
  library(scater)
  library(dplyr)
  library(reshape2)
  library(Matrix)
  library(purrr)
  library(scran)
  library(Seurat)
  library(tidyverse)
  library(here)
})
```

## color palettes
```{r col palettes}
colBRC <- c("#87c5af", "#94033b", "#fc8f3b", "#2580fe", "#514e72")
names(colBRC) <- c("LZFDC", "MRC", "TBRC","DZFDC","PRC2")

colLYM2 <- c("#a0b9c4", "#567a83", "#0C3C52", "#924440", "#eab562", "#b5a147")
names(colLYM2) <- c("naiveBcells", "GCBcells","plasmaCells","Tfh", "DCs",
                   "Myeloids")

colBoth <- c(colBRC, colLYM2) 
```


## load input data
```{r input data}
basedir <- here()
seuratLY <- readRDS(file = paste0(basedir, 
                              "/data/lympho/allSamples_mergedLab_seurat.rds"))
seuratBRC <- readRDS(file=paste0(basedir, "/data/folBRC_allSLO_seurat.rds"))

## group and sel for immune cells interacting with BRCs
selLab <- c("naiveBcells", "LZBcells", "DZshmBcells", "proDZ",
                   "plasmaCells", 
                   "Tfh", "DCs",
                   "mdMacrophages", "trMacrophages", "infMacrophages")
seuratLY <- subset(seuratLY, clusterLabel %in% selLab)
GCBs <- c("LZBcells", "DZshmBcells", "proDZ")
Myel <- c("mdMacrophages", "trMacrophages", "infMacrophages")
seuratLY$clusterLabel2 <- seuratLY$clusterLabel
seuratLY$clusterLabel2[which(seuratLY$clusterLabel %in% GCBs)] <- "GCBcells"
seuratLY$clusterLabel2[which(seuratLY$clusterLabel %in% Myel)] <- "Myeloids"

seuratLY$label_plus_SLO <- paste0(seuratLY$clusterLabel2, "_", seuratLY$SLO)
seuratLY$clusterLabel <- seuratLY$clusterLabel2


## subset on (naive) LN only
dim(seuratLY)
dim(seuratBRC)
seuratLY <- subset(seuratLY, SLO == "LN")
seuratBRC <- subset(seuratBRC, SLO == "LN" & cond=="naive")
#seuratBRC <- subset(seuratBRC, SLO == "LN")
dim(seuratLY)
dim(seuratBRC)

## plot input data
Idents(seuratLY) <- seuratLY$clusterLabel
DimPlot(seuratLY, reduction = "umap", cols = colLYM2)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

Idents(seuratBRC) <- seuratBRC$clusterLabel
DimPlot(seuratBRC, reduction = "umap", cols = colBRC)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


```

# merge BRCs and immune into one reference
```{r merge}

seuratBRC$group <- "BRC"
seuratLY$group <- "LYMPHO"


Idents(seuratLY) <- seuratLY$clusterLabel
Idents(seuratBRC) <- seuratBRC$clusterLabel

## merge
### fix assay and gene names
sceBRC <- as.SingleCellExperiment(seuratBRC)
sceLY <- as.SingleCellExperiment(seuratLY)

genesBRC <- data.frame(geneB=rownames(sceBRC)) %>% 
  mutate(ensID=gsub("\\..*","",geneB))
genesLY <- data.frame(geneL=rownames(sceLY)) %>% 
  mutate(ensID=gsub("\\..*","",geneL))
genesBoth <- full_join(genesBRC, genesLY, by="ensID") %>% 
  mutate(gene=ifelse(is.na(geneL), geneB, geneL)) %>% 
  dplyr::select(ensID, gene)
genesBRC <- genesBRC %>% left_join(., genesBoth, by="ensID")
genesLY <- genesLY %>% left_join(., genesBoth, by="ensID")

rownames(sceBRC) <- genesBRC$gene
rownames(sceLY) <- genesLY$gene

seuratBRC <- as.Seurat(sceBRC)
seuratLY <- as.Seurat(sceLY)

seuratBRC@assays$RNA <- seuratBRC@assays$originalexp
seuratLY@assays$originalexp <- seuratLY@assays$RNA

seuratMerge <- merge(seuratBRC, seuratLY)
seuratMerge <- ScaleData(seuratMerge, assay = "RNA")
seuratMerge <- ScaleData(seuratMerge, assay = "originalexp")

seuratMerge <- runSeurat3::rerunSeurat3(seuratMerge)

## vis merge
(table(seuratMerge$clusterLabel))

Idents(seuratMerge) <- seuratMerge$clusterLabel
DimPlot(seuratMerge, reduction = "umap", cols = colBoth)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```


## save reference
```{r save reference}

saveRDS(seuratMerge, file = paste0(basedir,
                                  "/data/LN_naiveBRCplusIMMmerged_seurat.rds"))
```


## session info
```{r session info}
sessionInfo()
date()
```



cd /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb
eval "$(pyenv init -)"
eval "$(pyenv virtualenv-init -)"

## run cellphonedb
### build database
cellphonedb database generate --user-interactions custom_interaction_file.csv

##run statistical analysis
## subsampling in R script.. 
cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_LN_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_LN_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outLN/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_PP_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_PP_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outPP/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100

cellphonedb method statistical_analysis /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_Spleen_metaData.txt /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_Spleen_countMatrix.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outSP/ --database /Users/mechthildluetge/Projects/scRNAseq/ILC_Hungwei/data/cellphonedb/out/cellphonedb_user_2021-10-05-09_17.db --debug-seed 3 --iterations 100


Count network interactions:
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_LN_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outLN/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outLN
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_PP_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outPP/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outPP
cellphonedb plot heatmap_plot /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/folBRC_LYMsub_Spleen_metaData.txt --pvalues-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outSP/pvalues.txt --output-path /Users/mechthildluetge/Projects/scRNAseq/CrossSLO_BRC_CXCL13/data/cellphonedb/outSP

We use CellPhoneDB as tool to predict cellular communication between SILT FRCs and ILC populations in our dataset. In brief, pairs of known interacting ligands and receptors are derived from public databases and cell types are analysed for enriched receptor–ligand interactions based on expression of a receptor by one cell type and a ligand by another cell type. 

deactivate  
