---
title: "organ specific signatures - BRCs only"
author: "Mechthild Lütge"
date: "14 May 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(SingleCellExperiment)
  library(RColorBrewer)
  library(viridis)
  library(ggsci)
  library(scater)
  library(scran)
  library(pheatmap)
  library(biomaRt)
  library(fgsea)
  library(grid)
  library(gridExtra)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(DOSE)
  library(enrichplot)
  library(msigdbr)
})

```

## plotting funct
```{r avg heatmap funct}

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  ## format gene names (depends on how gene list is inputed)
  selGenes <- selGenes %>% 
      mutate(geneID = (str_split(gene, '\\.', simplify = T)[,2]))
  #selGenes <- selGenes$labelNam
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>%
    filter(geneID %in% selGenes$geneID)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes$geneID,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}



```


## set dir and read input data
```{r set dir}
basedir <- here()

## read seurat objects with BRCs

seurat <- readRDS(paste0(basedir,"/data/folBRC_allSLO_seurat.rds"))
seurat <-rerunSeurat3(seurat)

```


## set color Vectors
```{r color vectors}

colPal <- c("#FF9933", "#660000","#CC6600", "#FF6666", "#389055")
names(colPal) <- c("LZFDC","MRC","DZFDC","TBRC", "PRC2")


colorSLO <- c(viridis(3)[c(1,2)], brewer.pal(n=9, name="YlOrRd")[c(4)])
colorCond <- brewer.pal(n=3, name="Paired")[c(1,2)]
colorSLOCond <- c("#440154FF","#807DBA", "#FEB24C", "#F16913", "#21908CFF")
colorBatch <- pal_igv()(n=length(unique(seurat$batch)))
colCl <- c(rcartocolor::carto_pal(name="Safe"),pal_aaas()(8))

names(colorSLO) <- c("LN", "PP", "Spleen")
names(colorCond) <- c("naive", "immunized")
names(colorBatch) <- unique(seurat$batch)
names(colorSLOCond) <- c("LN_immunized","LN_naive", "Spleen_naive",
                         "Spleen_immunized", "PP_immunized")

```

## visualize data {.tabset}

### clustering
```{r clustering}

DimPlot(seurat, reduction = "umap", cols=colPal, group.by = "clusterLabel")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```

### SLO
```{r SLO}

DimPlot(seurat, reduction = "umap", cols=colorSLO, group.by = "SLO")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seurat, reduction = "umap", cols=colorSLO, group.by = "SLO",
        pt.size=0.6)+
  theme_void()
```

### Batch
```{r Batch}

DimPlot(seurat, reduction = "umap", cols=colorBatch, group.by = "batch")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```

### Cond
```{r Cond}

DimPlot(seurat, reduction = "umap", cols=colorCond, group.by = "cond")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

```


### SLO plus cond
```{r SLO plus cond}

DimPlot(seurat, reduction = "umap", cols=colorSLOCond,group.by="SLO_plus_cond")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")


DimPlot(seurat, reduction = "umap",cols=colorSLOCond,group.by = "SLO_plus_cond",
        pt.size=0.6)+
  theme_void()
```


## vis fol BRC cluster characterization
```{r vis fol BRC cluster character, fig.height=5, fig.width=6}
seurat$SLO_clusterLabel <- paste0(seurat$SLO, "_", seurat$clusterLabel)
seurat$SLO_clusterLabel <- factor(seurat$SLO_clusterLabel, 
                                  levels = c("LN_LZFDC", "LN_MRC", "LN_DZFDC",
                                             "LN_TBRC", 
                                             "Spleen_LZFDC", "Spleen_MRC",
                                             "Spleen_DZFDC", "Spleen_TBRC",
                                             "PP_LZFDC", "PP_MRC", 
                                             "PP_DZFDC", "PP_TBRC"
                                             )) 

col_SLOlab <- c(colorLab, colorLab, colorLab)
names(col_SLOlab) <- c("LN_MRC","LN_TBRC","LN_LZFDC", "LN_DZFDC", 
                       "Spleen_MRC","Spleen_TBRC","Spleen_LZFDC","Spleen_DZFDC",
                      "PP_MRC", "PP_TBRC", "PP_LZFDC", "PP_DZFDC")


selMarker <- list(LZFDC = c("Fcamr", "Cr2", "Pthlh"),
                  MRC = c("Cxcl13", "Madcam1", "Tnfsf11"),
                  DZFDC = c("Pdlim3", "Crym", "Cxcl12"),
                  TBRC = c("Ccl21a","Ccl19", "Grem1")
                  )

genes <- data.frame(ensID=rownames(seurat)) %>%
  mutate(gene=gsub("^.*\\.", "", ensID))
selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub(".{1}$", "", grp)) %>% 
  left_join(., genes, by="gene") %>%
  filter(ensID != "ENSMUSG00000095320.Ccl21a")

Idents(seurat) <- seurat$SLO_clusterLabel

## sc heatmap across all
seuratSub <- subset(seurat, downsample = 250)
DoHeatmap(seuratSub, features = selGenes$ensID, group.by="SLO_clusterLabel",
          group.colors = col_SLOlab, angle = 90,
          draw.lines = F, size = 4, disp.min = -1)  +
  NoLegend() +
  scale_fill_continuous(type= "viridis")

## avg heatmap across all
selGenes$gene <- selGenes$ensID
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenes,
                  colVecIdent = colorSLO, 
                  ordVec=levels(seurat),
                  gapVecR=c(3,6,9), gapVecC=c(4,8),cc=F,
                  cr=F, condCol=T, colVecCond = colorLab)


```

### individual SLOs {.tabset}
```{r avgHeat ind SLO, warning=FALSE, echo = FALSE}

grpVec <- unique(seurat$SLO)

template_ah <- c(
    "#### {{grp}}\n",
    "```{r avgHeat ind SLO {{grp}}, echo = FALSE, fig.height=6, fig.width=5}\n",
    "seuratSub <- subset(seurat, SLO =='{{grp}}')",
    "pOut <- avgHeatmap(seurat = seuratSub, selGenes = selGenes,
                  colVecIdent = colorSLO, 
                  ordVec=levels(seuratSub),
                  gapVecR=c(3,6,9), gapVecC=NULL,cc=F,
                  cr=F, condCol=T, colVecCond = colorLab)\n",
    "```\n",
    "\n"
)
    
plots_ah <- lapply(grpVec, 
  function(grp) knitr::knit_expand(text = template_ah)
)

```

`r knitr::knit(text = unlist(plots_ah))`



## SLO specific BRC gene signatures
```{r SLO specific gene signatures}

Idents(seurat) <- seurat$SLO
seurat_markers_all <- FindAllMarkers(object = seurat,
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

write.table(seurat_markers_all,
            file=paste0(basedir, "/data/allBRC_SLOsignature.txt"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")




```

### vis top marker {.tabset}
#### across SLO
```{r vis top marker SLO, fig.height=9, fig.width=5}
selGenes <- seurat_markers_all %>% group_by(cluster) %>% 
  top_n(-15, p_val_adj) %>% 
  top_n(15, avg_log2FC)

Idents(seurat) <- seurat$SLO_plus_clusterlabel

pOut <- avgHeatmap(seurat = seurat, selGenes = selGenes,
                  colVecIdent = colorSLO, 
                  ordVec=levels(seurat),
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)

```


## SLO specific gene signatures per subset {.tabset}
```{r SLO specific gene signatures per subset}
seurat$clusterGrp <- seurat$clusterLabel

Idents(seurat) <- seurat$SLO
cwDEgenes <- lapply(unique(seurat$clusterGrp), function(grp){
  grpSub <- unique(seurat$clusterGrp)[which(unique(seurat$clusterGrp)==grp)]
  seuratSub <- subset(seurat, clusterGrp == grpSub)
  DEgenes <-FindAllMarkers(seuratSub, only.pos=T, logfc.threshold = 0.25,
                           min.pct = 0.1)
  DEgenes <- DEgenes %>% filter(p_val_adj<0.01)
})

names(cwDEgenes) <- unique(seurat$clusterGrp)

```



```{r avgHeat cwDE, warning=FALSE, echo = FALSE}

grpVec <- unique(seurat$clusterGrp)
seurat$clusterGrp_plus_SLO <- as.factor(paste0(seurat$clusterGrp,
                                              "_",seurat$SLO))
Idents(seurat) <- seurat$clusterGrp_plus_SLO

template_ah <- c(
    "### {{grp}}\n",
    "```{r avgHeat cwDE {{grp}}, echo = FALSE, fig.height=8, fig.width=6}\n",
    "DeSub <- cwDEgenes[['{{grp}}']]",
    "DeSel <- DeSub %>% group_by(cluster) %>% top_n(., 10, avg_logFC)",
    "gapVecCol <- seq(3, length(levels(seurat$clusterGrp_plus_SLO)), by=3)",
    "pOut <- avgHeatmap(seurat = seurat, selGenes = DeSel,
                     colVecIdent = colClGrp, colVecCond=colorSLO,
                     ordVec=levels(seurat$clusterGrp_plus_SLO),
                     gapVecR=c(10,20), gapVecC=gapVecCol,cc=FALSE,
                     cr=F, condCol=T)\n",
    "```\n",
    "\n"
)
    
plots_ah <- lapply(grpVec, 
  function(grp) knitr::knit_expand(text = template_ah)
)

```

`r knitr::knit(text = unlist(plots_ah))`


## SLO marker overlap between grps 
```{r SLO marker overlap, fig.height=21, fig.width=9}

grpVec <- unique(seurat$clusterGrp)
orgVec <- unique(seurat$SLO)

orgMarker <- lapply(orgVec, 
  function(slo){
    TBRCgenes <- cwDEgenes[["TBRC"]] %>% filter(cluster==slo)
    MRCgenes <- cwDEgenes[["MRC"]] %>% filter(cluster==slo)
    LZFDCgenes <- cwDEgenes[["LZFDC"]] %>% filter(cluster==slo)
    DZFDCgenes <- cwDEgenes[["DZFDC"]] %>% filter(cluster==slo)
    all <- intersect(intersect(intersect(TBRCgenes$gene, MRCgenes$gene),
                               LZFDCgenes$gene), DZFDCgenes$gene)
  }
)

names(orgMarker) <- orgVec
## orgMarker = DE genes in all shared folicular BRC subsets
## (adj_pVal < 0.01, min_pct = 0.1, log_FC > 0.25)
```

## vis all marker conserved across all subsets {.tabset}
### LN
```{r vis all conserved marker LN, fig.height=16, fig.width=12}

seurat$label_plus_organ_plus_cond <- factor(paste0(seurat$SLO_plus_clusterlabel, "_",
                                            seurat$cond))
Idents(seurat) <- seurat$label_plus_organ_plus_cond
selGenes <- data.frame(gene=orgMarker[["LN"]])
allGenes <- data.frame(gene=rownames(seurat)) %>%
  mutate(geneNam=gsub("^.*\\.", "", gene))
selGenesDat <- selGenes %>% left_join(., allGenes, by=c("gene"))

seurat$label_plus_cond <- paste0(seurat$clusterLabel, "_", seurat$cond)
colSel <- c(rep(colorCond[["naive"]], 5), rep(colorCond[["immunized"]], 5))
names(colSel) <- unique(seurat$label_plus_cond)
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesDat,
                  colVecIdent = colorSLO, 
                  ordVec=levels(seurat),
                  gapVecC=c(10,15),cc=F,
                  cr=T,condCol=T, colVecCond=colSel)
```

### Spleen
```{r vis all conserved marker spleen, fig.height=12, fig.width=12}

selGenes <- data.frame(gene=orgMarker[["Spleen"]])
allGenes <- data.frame(gene=rownames(seurat)) %>%
  mutate(geneNam=gsub("^.*\\.", "", gene))
selGenesDat <- selGenes %>% left_join(., allGenes, by=c("gene")) %>% 
  filter(nchar(geneNam)>1)

pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesDat,
                  colVecIdent = colorSLO, 
                  ordVec=levels(seurat),
                  gapVecC=c(10,15),cc=F,
                  cr=T,condCol=T, colVecCond=colSel)
```

### PP
```{r vis all conserved marker PP, fig.height=8, fig.width=12}

selGenes <- data.frame(gene=orgMarker[["PP"]])
allGenes <- data.frame(gene=rownames(seurat)) %>%
  mutate(geneNam=gsub("^.*\\.", "", gene))
selGenesDat <- selGenes %>% left_join(., allGenes, by=c("gene"))

pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesDat,
                  colVecIdent = colorSLO, 
                  ordVec=levels(seurat),
                  gapVecC=c(10,15),cc=F,
                  cr=T,condCol=T, colVecCond=colSel)
```



## GSEA clusterProfiler {.tabset}
### LN
```{r GSEA clusterprofiler, fig.height=6, fig.width=12}
orgMarkerDat <- data.frame(do.call("cbind", orgMarker)) %>% 
  gather(., SLO, gene, LN:Spleen) %>% mutate(EnsID = gsub("\\..*$", "", gene))

orgMarkerDatLN <- orgMarkerDat %>% filter(SLO == "LN")
egoLN <- enrichGO(gene         = orgMarkerDatLN$EnsID,
                OrgDb         = org.Mm.eg.db,
                keyType       = 'ENSEMBL',
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05)
egoLN <- setReadable(egoLN, OrgDb = org.Mm.eg.db)
dotplot(egoLN, showCategory=30)
cnetplot(egoLN, node_label="all") 
emapplot(egoLN, showCategory = 15)


orgMarkerDatLN <- bitr(orgMarkerDatLN$EnsID, fromType = "ENSEMBL",
        toType = c("ENTREZID"),
        OrgDb = org.Mm.eg.db)
eKeggLN <- enrichKEGG(gene    = orgMarkerDatLN$ENTREZID,
                 organism     = 'mmu',
                 pvalueCutoff = 0.05)
dotplot(eKeggLN, showCategory=30)
cnetplot(eKeggLN, node_label="all") 
emapplot(eKeggLN, showCategory = 15)


m_t2g <- msigdbr(species = "Mus musculus", category = "H") %>% 
  dplyr::select(gs_name, entrez_gene)
em <- enricher(orgMarkerDatLN$ENTREZID, TERM2GENE=m_t2g, minGSSize=5,
               pvalueCutoff = 0.5, qvalueCutoff = 0.5)
dotplot(em, showCategory=30)
cnetplot(em, node_label="all") 
emapplot(em, showCategory = 15)

```


### PP
```{r GSEA clusterprofiler PP, fig.height=8, fig.width=14}

orgMarkerDatPP <- orgMarkerDat %>% filter(SLO == "PP")
egoPP <- enrichGO(gene         = orgMarkerDatPP$EnsID,
                OrgDb         = org.Mm.eg.db,
                keyType       = 'ENSEMBL',
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05)
egoPP <- setReadable(egoPP, OrgDb = org.Mm.eg.db)
dotplot(egoPP, showCategory=30)
cnetplot(egoPP, node_label="all") 
emapplot(egoPP, showCategory = 15)

orgMarkerDatPP <- bitr(orgMarkerDatPP$EnsID, fromType = "ENSEMBL",
        toType = c("ENTREZID"),
        OrgDb = org.Mm.eg.db)
eKeggPP <- enrichKEGG(gene    = orgMarkerDatPP$ENTREZID,
                 organism     = 'mmu',
                 pvalueCutoff = 0.1)
dotplot(eKeggPP, showCategory=30)
emapplot(eKeggPP, showCategory = 15)

em <- enricher(orgMarkerDatPP$ENTREZID,
               TERM2GENE=m_t2g,
               minGSSize=5,
               pvalueCutoff = 0.5,
               qvalueCutoff = 0.5)
dotplot(em, showCategory=30)
cnetplot(em, node_label="all") 
emapplot(em, showCategory = 15)

```

### Spleen
```{r GSEA clusterprofiler Spleen, fig.height=8, fig.width=12}
 
orgMarkerDatSP <- orgMarkerDat %>% filter(SLO == "Spleen")
egoSP <- enrichGO(gene         = orgMarkerDatSP$EnsID,
                OrgDb         = org.Mm.eg.db,
                keyType       = 'ENSEMBL',
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05)
egoSP <- setReadable(egoSP, OrgDb = org.Mm.eg.db)
dotplot(egoSP, showCategory=30)
cnetplot(egoSP, node_label="all") 
emapplot(egoSP, showCategory = 15)

orgMarkerDatSP <- bitr(orgMarkerDatSP$EnsID, fromType = "ENSEMBL",
        toType = c("ENTREZID"),
        OrgDb = org.Mm.eg.db)
eKeggSP <- enrichKEGG(gene    = orgMarkerDatSP$ENTREZID,
                 organism     = 'mmu',
                 pvalueCutoff = 0.1)
dotplot(eKeggSP, showCategory=30)
cnetplot(eKeggSP, node_label="all")
emapplot(eKeggSP, showCategory = 15)

em <- enricher(orgMarkerDatSP$ENTREZID,
               TERM2GENE=m_t2g,
               minGSSize=5,
               pvalueCutoff = 0.5,
               qvalueCutoff = 0.5)
dotplot(em, showCategory=30)
cnetplot(em, node_label="all") 
emapplot(em, showCategory = 15)

```


### vis summary all
```{r vis summary, fig.height=6, fig.width=9}
selLN <- c("GO:0010463", "GO:0071604", "GO:0048706", "GO:2000146", "GO:0006959",
           "GO:1901342")
selPW_LN <- egoLN@result %>% filter(p.adjust<0.05) %>% 
  filter(ID %in% selLN) %>%
  mutate(IDfac= factor(ID, levels = selLN)) %>% 
  arrange(., IDfac) %>% 
  dplyr::select(ID, Description, GeneRatio, p.adjust, Count) %>% 
  mutate(SLO="LN")

selSP <- c("GO:0019882", "GO:0048536", "GO:0098656", "GO:0008015", "GO:0035265",
           "GO:0060537")
selPW_SP <- egoSP@result %>% filter(p.adjust<0.05) %>% 
  filter(ID %in% selSP) %>% 
  mutate(IDfac= factor(ID, levels = selSP)) %>% 
  arrange(., IDfac) %>% 
  dplyr::select(ID, Description, GeneRatio, p.adjust, Count) %>% 
  mutate(SLO="Spleen")

selPP <- c("GO:0030509", "GO:0016331","GO:0030198", "GO:0061448", "GO:0042573",
           "GO:0007178")
selPW_PP <- egoPP@result %>% filter(p.adjust<0.05) %>% 
  filter(ID %in% selPP) %>% 
  mutate(IDfac= factor(ID, levels = selPP)) %>% 
  arrange(., IDfac) %>% 
  dplyr::select(ID, Description, GeneRatio, p.adjust, Count) %>% 
  mutate(SLO="PP")


selPW_all <- rbind(selPW_PP, selPW_SP, selPW_LN) %>%
  mutate(nom=as.numeric(gsub("\\/.*", "", GeneRatio))) %>% 
  mutate(denom=as.numeric(gsub(".*\\/", "", GeneRatio))) %>% 
  mutate(geneRatio=nom/denom) %>% 
  mutate(one_minPadjust=(1-p.adjust))
 
level_order <- unique(selPW_all$Description)
ggplot(selPW_all, aes(x=factor(Description, level= level_order), y=geneRatio)) + 
  geom_point(aes(col=SLO, size=Count, alpha=one_minPadjust)) + 
  scale_color_manual(values = colorSLO)+
  scale_alpha_continuous(range = c(0.5,1), name="p.adjust",
                         labels=c(0.04, 0.03, 0.02, 0.01),
                         breaks = c(0.96, 0.97, 0.98, 0.99)) +
  scale_size_continuous(range = c(3,6), labels = c(2,4,6,8),
                        breaks = c(2,4,6,8)) +
  geom_segment(aes(x=Description, 
                   xend=Description, 
                   y=min(geneRatio), 
                   yend=max(geneRatio)), 
               linetype="dashed", 
               size=0.1) +   # Draw dashed lines
  coord_flip() +
  theme_minimal() +
  theme(panel.border = element_rect(fill = NA), 
        panel.grid = element_blank(), axis.title.y = element_blank())
```



## vis sel marker across all subsets {.tabset}
### LN
```{r vis sel marker LN, fig.height=7, fig.width=6}
selGenes <- data.frame(geneNam = c("Ebf1", "Ebf3", "Hoxd9", "Ntrk1", "Egr3",
                                   "Sfrp2", "Adh1","Gata6", "Lhx8",
                                   "Scara5", "Ptn", "Inmt",
                                   "Cldn1", "Adam12",
                                   "Cxcl1", "Ccl11"))
allGenes <- data.frame(gene=rownames(seurat)) %>%
  mutate(geneNam=gsub("^.*\\.", "", gene))
selGenesDat <- selGenes %>% left_join(., allGenes, by=c("geneNam"))

Idents(seurat) <- seurat$label_plus_organ
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesDat,
                  colVecIdent = colorLab, 
                  ordVec=levels(seurat),
                  gapVecR=c(9,12,14), gapVecC=c(4,8),cc=F,
                  cr=F, condCol=T, colVecCond=colorSLO)

```

### Spleen
```{r vis sel marker Spleen, fig.height=6, fig.width=5}
selGenes <- data.frame(geneNam = c("Tlx1", "Coch", "Ttn", "Pcdh10", "Gabra2",
                                   "Gal3st2", "Lyz2","Cr2"))
allGenes <- data.frame(gene=rownames(seurat)) %>%
  mutate(geneNam=gsub("^.*\\.", "", gene))
selGenesDat <- selGenes %>% left_join(., allGenes, by=c("geneNam"))

Idents(seurat) <- seurat$label_plus_organ
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesDat,
                  colVecIdent = colorLab, 
                  ordVec=levels(seurat),
                  gapVecR=c(1,3,7), gapVecC=c(4,8),cc=F,
                  cr=F, condCol=T, colVecCond=colorSLO)

```


### PP
```{r vis sel marker PP, fig.height=6, fig.width=5}
selGenes <- data.frame(geneNam = c("Mab21l2", "Dkk2", "Mgp", "Lum", "Mfap4",
                                   "Sytl2", "Musk","Spon2", "Cxcl14"))
allGenes <- data.frame(gene=rownames(seurat)) %>%
  mutate(geneNam=gsub("^.*\\.", "", gene))
selGenesDat <- selGenes %>% left_join(., allGenes, by=c("geneNam"))

Idents(seurat) <- seurat$label_plus_organ
pOut <- avgHeatmap(seurat = seurat, selGenes = selGenesDat,
                  colVecIdent = colorLab, 
                  ordVec=levels(seurat),
                  gapVecR=c(3,5,8), gapVecC=c(4,8),cc=F,
                  cr=F, condCol=T, colVecCond=colorSLO)

```


## organ specific genes dotplot
```{r SLO sepcific genes, fig.height=8, fig.width=7}

selMarkerAll <- read_tsv(paste0(basedir,"/data/SLOspecificSelMarker.txt"))


genes <- data.frame(ensID=rownames(seurat)) %>%
  mutate(gene=gsub("^.*\\.", "", ensID))
selGenes <- selMarkerAll %>% 
  left_join(., genes, by=c("geneNam"="gene"))

Idents(seurat) <- seurat$SLO_clusterLabel
clusterDat <- data.frame(ident=levels(seurat)) %>%
  mutate(SLO=gsub("_.*$", "", ident)) %>% 
  mutate(namFull=paste0(ident, "_", SLO)) %>%
  mutate(label= gsub("^.*_", "", ident))

DotPlot(seurat, assay="RNA", features = selGenes$ensID, split.by = "SLO",
        cols = colorSLO[c(1,3,2)], col.max = 6) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$ensID, labels=selGenes$geneNam) +
  scale_y_discrete(breaks=clusterDat$namFull, labels=clusterDat$label) +
  xlab("") + ylab("")


DotPlot(seurat, assay="RNA", features = selGenes$ensID, split.by = "SLO",
        cols = colorSLO[c(1,3,2)], col.max = 6) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=selGenes$ensID, labels=selGenes$geneNam) +
  scale_y_discrete(breaks=clusterDat$namFull, labels=clusterDat$label) +
  xlab("") + ylab("") + theme(legend.position = "none")

```



## save object
```{r save seurat}
saveRDS(seurat, file=paste0(basedir, "/data/allBRC_seurat.rds"))
```


## session info
```{r session info}
sessionInfo()
date()
```

