---
title: "vis merged samples May20"
author: "Mechthild Lütge"
date: "14 May 2020"
output:
  html_document:
    self_contained: no
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r libs}
suppressPackageStartupMessages({
  library(tidyverse)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(here)
  library(runSeurat3)
  library(SingleCellExperiment)
  library(RColorBrewer)
  library(viridis)
  library(ggsci)
  library(scater)
  library(scran)
  library(pheatmap)
  library(CellMixS)
  library(muscat)
})

```


## set dir and read input data
```{r set dir}
basedir <- here()
sce <- readRDS(paste0(basedir, "/data/allSamplesMerged_sce.rds"))
sce$SLO_plus_cond <- paste0(sce$SLO, '_' , sce$cond)
dataCnts <- data.frame(table(sce$dataset))
colnames(dataCnts) <- c('dataset', 'cnt_all')


```


## set color Vectors
```{r color vectors}

colorSLO <- c(viridis(3)[c(1,2)], brewer.pal(n=9, name="YlOrRd")[c(4)])
colorCond <- brewer.pal(n=3, name="Paired")[c(1,2)]
colorSLOCond <- c("#440154FF","#807DBA", "#FEB24C", "#F16913", "#21908CFF")
colorBatch <- pal_igv()(n=length(unique(sce$batch)))

# set colorPal names
names(colorSLO) <- c("LN", "PP", "Spleen")
names(colorCond) <- c("naive", "immunized")
names(colorBatch) <- unique(sce$batch)
names(colorSLOCond) <- c("LN_immunized","LN_naive", "Spleen_naive",
                         "Spleen_immunized", "PP_immunized")


```


## keep only Cxcl13 positive cells
```{r only cxcl13 expr cells}
## keep only eyfp expressing cells
sce <- sce[,which(logcounts(sce)[
  which(rownames(sce)=='Rosa26eyfp.Rosa26eyfp'),] > 0)]
dataCnts_post <- data.frame(table(sce$dataset))
colnames(dataCnts_post) <- c('dataset', 'cnt_FRC')

dataCnts <- dataCnts %>% left_join(., dataCnts_post, by = 'dataset')

## keep only cxcl13 expressing cells
sce2 <- sce[,which(logcounts(sce)[
  which(rownames(sce)=='ENSMUSG00000023078.Cxcl13'),] > 0)]
dataCnts_post <- data.frame(table(sce2$dataset))
colnames(dataCnts_post) <- c('dataset', 'cnt_BRC')

dataCnts <- dataCnts %>% left_join(., dataCnts_post, by = 'dataset')

remove(sce2)
```


## plot data summaries {.tabset}

```{r data summaries}

metaDat <- data.frame(colData(sce)) %>% 
  select(SLO, dataset, batch, cond, SLO_plus_cond) %>% distinct()

dataCnts <- dataCnts %>% left_join(., metaDat, by = 'dataset')
```

### color SLO plus cond
```{r cnt per SLO plus cond}

ggplot(dataCnts, aes(x = SLO_plus_cond, y = cnt_FRC)) +
  geom_bar(aes(color = SLO_plus_cond, fill = SLO_plus_cond, alpha = 0.8),
           stat = "identity") +
  scale_color_manual(values = colorSLOCond) +
  scale_fill_manual(values = colorSLOCond) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```


### color BRC
```{r BRC cnt per SLO plus cond}

dataCnts <- dataCnts %>% mutate(cnt_noBRC=cnt_FRC-cnt_BRC)
dataCnts_long <- gather(dataCnts, BRC, cnt, cnt_BRC, cnt_noBRC) %>% 
  mutate(BRClab = ifelse(BRC=="cnt_BRC", "Cxcl13_pos", "Cxcl13_neg"))

ggplot(dataCnts_long, aes(x = SLO_plus_cond, y = cnt)) +
  geom_bar(aes(color = BRClab, fill = BRClab),
           stat = "identity") +
  scale_color_manual(values = c("#A0A0A0", "#009999")) +
  scale_fill_manual(values = c("#A0A0A0", "#009999")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


### color batches
```{r batch summary}

ggplot(dataCnts, aes(x = SLO_plus_cond, y = cnt_FRC)) +
  geom_bar(aes(color = batch, fill = batch, alpha = 0.9),
           stat = "identity") +
  scale_color_manual(values = colorBatch) +
  scale_fill_manual(values = colorBatch) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```

## DR and clustering

### variance components per sample {.tabset}
```{r var comp per smp}

datSet <- unique(sce$dataset)
sce_list <- lapply(datSet, function(dat){
  sceSub <- sce[,sce$dataset == dat]
})
names(sce_list) <- datSet

#determine variance components for each sce separately
dec_list <- lapply(sce_list, function(sceSub){
  dec <- modelGeneVar(sceSub)
})

names(dec_list) <- names(sce_list)

#Get hvgs
hvg_list <- lapply(dec_list, function(dec){
  hvg <- getTopHVGs(dec, n=1000)
})

names(hvg_list) <- names(sce_list)

# finding common hvg
combined.dec <- do.call(combineVar, dec_list)
chosen.hvgs <- combined.dec$bio > 0
sum(chosen.hvgs)

```


```{r varPlot}

# plot variance
template_vp <- c(
    "#### {{dat}}\n",
    "```{r varPlot {{dat}}, echo = FALSE}\n",
    "dec <- dec_list[['{{dat}}']]",
    "fit <- metadata(dec)",
    "plot(fit$mean, fit$var, xlab='Mean of log-expression',
      ylab='Variance of log-expression')",
    "curve(fit$trend(x), col='dodgerblue', add=TRUE, lwd=2)\n",
    "```\n",
    "\n"
)

plots_vp <- lapply(names(dec_list), 
  function(dat) knitr::knit_expand(text = template_vp)
)

```

`r knitr::knit(text = unlist(plots_vp))`


```{r DR and clustering}

sce <- runPCA(sce, subset_row=chosen.hvgs,
    BSPARAM=BiocSingular::RandomParam(), ncomponents = 30)
sce <- runTSNE(sce, dimred="PCA", perplexity=20)
sce <- runUMAP(sce, subset_row=chosen.hvgs, dimred="PCA", pca =30,
               spread = 2, min_dist = 0.1, n_neighbors=50, ntop = 5000)
  
#clustering 
#g <- buildSNNGraph(sce, k=50, use.dimred="PCA")
#clusters <- igraph::cluster_walktrap(g)$membership

```

## vis Data {.tabset}
### SLO {.tabset}
#### UMAP
```{r SLO umap}

p1 <- visGroup(sce, group = "SLO", dim_red = "UMAP") +
  scale_color_manual(values = colorSLO) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

p1


```

#### TSNE
```{r SLO tsne}

p1 <- visGroup(sce, group = "SLO", dim_red = "TSNE") +
  scale_color_manual(values = colorSLO) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("TSNE1") +
  ylab("TSNE2")

p1


```


### COND {.tabset}
#### UMAP
```{r COND umap}

p1 <- visGroup(sce, group = "cond", dim_red = "UMAP") +
  scale_color_manual(values = colorCond) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

p1


```

#### TSNE
```{r COND tsne}

p1 <- visGroup(sce, group = "cond", dim_red = "TSNE") +
  scale_color_manual(values = colorCond) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("TSNE1") +
  ylab("TSNE2")

p1


```


### BATCH {.tabset}
#### UMAP
```{r Batch umap}

p1 <- visGroup(sce, group = "batch", dim_red = "UMAP") +
  scale_color_manual(values = colorBatch) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

p1


```

#### TSNE
```{r BATCH tsne}

p1 <- visGroup(sce, group = "batch", dim_red = "TSNE") +
  scale_color_manual(values = colorBatch) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("TSNE1") +
  ylab("TSNE2")

p1


```

### SLO plus COND {.tabset}
#### UMAP
```{r SLO_plus_cond umap}

p1 <- visGroup(sce, group = "SLO_plus_cond", dim_red = "UMAP") +
  scale_color_manual(values = colorSLOCond) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

p1


```

#### TSNE
```{r SLO_plus_cond tsne}

p1 <- visGroup(sce, group = "SLO_plus_cond", dim_red = "TSNE") +
  scale_color_manual(values = colorSLOCond) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("TSNE1") +
  ylab("TSNE2")

p1


```


## save object
```{r save object}

saveRDS(sce, paste0(basedir, "/data/allSamplesMerged_plusDR_sce.rds"))

```


## MDS plot 

```{r mds plot, fig.height=3, fig.width=4}
sce <- prepSCE(sce, 
    kid = "SLO", 
    gid = "cond",   
    sid = "dataset",  
    drop = F)       

pb <- aggregateData(sce,
    assay = "counts", fun = "sum",
    by = c("cluster_id", "sample_id"))

pb_mds <- pbMDS(pb) +
  scale_color_manual(values = colorSLO)
pb_mds

```



## session info
```{r session info}
sessionInfo()
date()
```

